Figures, Plots, Graphs

### 2025-03-01 notebooks/02-tables.qmd
checkpoint

# quick, informative, easy table of sample demographics w/ some other questions.
tbl1 <- data |>  
  labelled::set_variable_labels(
    age_cat = "Age group, in years"
  ) |> 
  gtsummary::tbl_summary(
    by = qset,
    include = c(
    qset,
    age_cat, 
    gender_3cat, 
    race, 
    educ_4cat, 
    partyid_3cat,
    ideo,
    ideolean,
    voted2020,
    choice2020, 
    voteintent,
    milserv1,
    milserv2,
    milservfam),
    statistic = list(gtsummary::all_categorical()~ "{n} ({p}%)"),
    type = list(milserv1 ~ "categorical",

                milserv2 ~ "categorical",
                milservfam ~"categorical"), 
    missing = "no") |> 
  gtsummary::add_n() |> 
  gtsummary::add_overall() |> 
  gtsummary::modify_header(label = "**Variable**") |> 
  gtsummary::bold_labels()

gtsummary::as_gt(tbl1) |>  
  gt::tab_footnote(
    footnote = "'true' Independents do not identify nor 'lean' toward either political party",
    placement = "left"
    ) |>  
  gt::cols_align(align = 'left', columns = everything()) |> 
  gt::tab_options(
    table.font.size = "small",
    data_row.padding = gt::px(1)
  )

@@

# Likert table summary of Trust and confidence variables
tbl_tc <- data |> 
  select(q19:q24) |>
  gtsummary::tbl_likert(
    statistic = ~"{n} ({p}%)", 
    label = list(q19 = "Votes will be counted as intended",
                 q20 = "Election officials will do good job",
                 q21 = "Election workers will be committed",
                 q22 = "Voting process will be fair",
                 q23 = "Voting outcome will be fair",
                 q24 = "Election technology will be secure"))

# add continuous summary of Likert scores. Merge tables into one
tbl_tc_summary <- list(
  tbl_tc,
  gtsummary::tbl_wide_summary(
    data |>
      select(q19:q24) |> mutate(across(where(is.factor), as.numeric)),
    statistic = c("{mean}", "{sd}"),
    label = list(
      q19 = "Votes will be counted as intended",
      q20 = "Election officials will do good job",
      q21 = "Election workers will be committed",
      q22 = "Voting process will be fair",
      q23 = "Voting outcome will be fair",
      q24 = "Election technology will be secure"
    ),
    type = ~ "continuous"
  )
) |>
  gtsummary::tbl_merge(tab_spanner = F) |> 
  gtsummary::modify_header(label="")

# print as huxtable and modify using `huxtable` R package
gtsummary::as_hux_table(tbl_tc_summary) |>
  huxtable::set_width(0.6) |> 
  huxtable::set_col_width(rep(c(0.4))) |> 
  huxtable::theme_article() |> 
  huxtable::set_position("center")

>>>>

# quick, informative, easy table of sample demographics w/ some other questions.

# best method for html document
tbl.desc <- data |>  
  labelled::set_variable_labels(
    age_4cat = "Age group, in years",
    gender_3cat = "Gender",
    race2 = "Race/Ethnicity",
    educ_4cat = "Education",
    partyid_3cat = "Party ID"
  ) |> 
  gtsummary::tbl_summary(
    by = group,
    include = c(
    age_4cat, 
    gender_3cat, 
    race2, 
    educ_4cat, 
    partyid_3cat),
    statistic = list(gtsummary::all_categorical()~ "{n} ({p}%)"), 
    missing = "ifany",
    missing_text = "NA") |> 
  gtsummary::add_n() |> 
  gtsummary::add_overall() |> 
  gtsummary::modify_header(label = "**Variable**") |> 
  gtsummary::bold_labels()

gtsummary::as_gt(tbl.desc) |>  
  gt::tab_footnote(
    footnote = "Independents do not identify nor 'lean' toward either political party",
    placement = "left"
    ) |>  
  gt::cols_align(align = 'left', columns = everything()) |> 
  gt::tab_options(
    table.font.size = "small",
    data_row.padding = gt::px(1)
  )

gtsummary::as_kable_extra(
  tbl.desc, 
  booktabs = T,
  digits = 2,
  addtl_fmt = T,
  linesep = ""
) |> 
  kableExtra::kable_styling(
    latex_options = c("striped", "scale_down"),
    font_size = 8)

@@

# Likert table summary of Trust and confidence variables
# best method for html document
tbl.desc <- data |>  
  labelled::set_variable_labels(
    age_4cat = "Age group, in years",
    gender_3cat = "Gender",
    race2 = "Race/Ethnicity",
    educ_4cat = "Education",
    partyid_3cat = "Party ID"
  ) |> 
  gtsummary::tbl_summary(
    by = group,
    include = c(
    age_4cat, 
    gender_3cat, 
    race2, 
    educ_4cat, 
    partyid_3cat),
    statistic = list(gtsummary::all_categorical()~ "{n} ({p}%)"), 
    missing = "ifany",
    missing_text = "NA") |> 
  gtsummary::add_n() |> 
  gtsummary::add_overall() |> 
  gtsummary::modify_header(label = "**Variable**") |> 
  gtsummary::bold_labels()

gtsummary::as_gt(tbl.desc) |>  
  gt::tab_footnote(
    footnote = "Independents do not identify nor 'lean' toward either political party",
    placement = "left"
    ) |>  
  gt::cols_align(align = 'left', columns = everything()) |> 
  gt::tab_options(
    table.font.size = "small",
    data_row.padding = gt::px(1)
  )

gtsummary::as_kable_extra(
  tbl.desc, 
  booktabs = T,
  digits = 2,
  addtl_fmt = T,
  linesep = ""
) |> 
  kableExtra::kable_styling(
    latex_options = c("striped", "scale_down"),
    font_size = 8)

### 2025-03-01 notebooks/03-figures.qmd
checkpoint

>>>>

```{r}
#| label: fig-itemcor
#| fig-cap: "Inter-item correlations between AZ and local area items"

# Inter-item correlations between AZ and Local items
# use the `easystats` correlation package
item.cor <- df |>
  dplyr::rename(
    trustvote_az     = q19,
    trustadmin_az    = q20,
    truststaff_az    = q21,
    trustprcss_az    = q22,
    trusttech_az     = q24,
    votefraud_az     = q28_1,
    miscount_az      = q28_2,
    turnedaway_az    = q28_3,
    foreigntamp_az   = q28_4,
    discourage_az    = q28_5,
    trustvote_lcl    = q30,
    trustadmin_lcl   = q31,
    truststaff_lcl   = q32,
    trustprcss_lcl   = q33,
    trusttech_lcl    = q35,
    votefraud_lcl    = q40_1,
    miscount_lcl     = q40_2,
    turnedaway_lcl   = q40_3,
    foreigntamp_lcl  = q40_4,
    discourage_lcl   = q40_5
  ) |> 
  correlation::correlation(
    select = 
      c("trustvote_az", "trustadmin_az", "truststaff_az", "trustprcss_az", "trusttech_az",
        "votefraud_az", "miscount_az", "turnedaway_az", "foreigntamp_az", "discourage_az"),
    select2 =
      c("trustvote_lcl", "trustadmin_lcl", "truststaff_lcl", "trustprcss_lcl","trusttech_lcl",
        "votefraud_lcl", "miscount_lcl", "turnedaway_lcl", "foreigntamp_lcl","discourage_lcl"),
    method = "polychoric") |>
  summary()

layers <- correlation::visualisation_recipe(
  item.cor, labs = list(title = "Polychoric Item Correlation Matrix"))
layers

plot(layers) + theme(axis.text.x = element_text(angle = 45, vjust = 0.6))

# Likert Plots

```{r}
#| label: fig-trustlikert
#| fig-cap: "Trust in Election Administration by Experiment Condition"
#| fig-height: 4.5
#| fig-pos: "h"

# Likert plot of trust and confidence in election admin comparing treatment and
# control groups
trust.likert.treat.az <- df |>
  # relabel factor labels for plotting
  mutate(across(
    c(q19, q20, q21, q22, q23, q24, q30, q31, q32, q33, q34, q35),
    ~ pewmethods::fct_case_when(
      .x %in% c("Not at all confident", "Not at all committed") ~ "Not at all confident",
      .x %in% c("Not too confident", "Not too committed") ~ "Not too confident",
      .x %in% c("Somewhat confident", "Somewhat committed") ~ "Somewhat confident",
      .x %in% c("Very confident", "Very committed") ~ "Very confident",
      TRUE ~ NA
    )
  )) |>
  ggstats::gglikert(
    include = c(q19, q20, q21, q22, q24),
    y = 'group',
    facet_rows = vars(.question),
    symmetric = F,
    labels_size = 2.5,
    add_totals = T
  ) +
  ggplot2::scale_fill_grey(start = 0.5, end = 0.1)+
  # ggplot2::scale_fill_brewer(palette = "Greys")+
  labs(subtitle = str_wrap("Maricopa County, Arizona", width = 95)) +
  theme_bw() +
  theme(
    text = element_text(family = "Palatino Linotype"),
    legend.position = 'bottom',
    axis.text.x = element_blank(), # remove percentage text along x-axis
    strip.text.y.right = element_blank()
  )

# confidence in elections between experiment conditions for items concerning local area
trust.likert.treat.lcl <- df |>
  # relabel factor labels for plotting
  mutate(across(
    c(q19, q20, q21, q22, q23, q24, q30, q31, q32, q33, q34, q35),
    ~ pewmethods::fct_case_when(
      .x %in% c("Not at all confident", "Not at all committed") ~ "Not at all confident",
      .x %in% c("Not too confident", "Not too committed") ~ "Not too confident",
      .x %in% c("Somewhat confident", "Somewhat committed") ~ "Somewhat confident",
      .x %in% c("Very confident", "Very committed") ~ "Very confident",
      TRUE ~ NA
    )
  )) |>
  ggstats::gglikert(
    include = c(q30, q31, q32, q33, q35),
    y = 'group',
    facet_rows = vars(.question),
    variable_labels = c(
      q30 = "Votes counted as intended",
      q31 = "Election officials will do good job",
      q32 = "Election workers will be committed",
      q33 = "Voting process will be fair",
      q35 = "Election technology will be secure"
    ),
    facet_label_wrap = 10,
    labels_size = 2.5,
    symmetric = F,
    add_totals = T
  ) +
  ggplot2::scale_fill_grey(start = 0.5, end = 0.1)+
  # ggplot2::scale_fill_brewer(palette = "Greys")+
  labs(subtitle = str_wrap("Local Area", width = 95)) +
  theme_bw() +
  theme(
    text = element_text(family = "Palatino Linotype"),
    legend.position = 'bottom',
    axis.text.y.left = element_blank(), # remove y-axis text on 2nd plot
    axis.text.x = element_blank(), # remove percentage text along x-axis
    strip.text.y.right = element_text(angle = 0, size = 8)
  )

# trust.likert.treat.az
# trust.likert.treat.lcl

# Instead of two separate plots, arrange into one plot while remaining separated
# use `ggpubr` package to print multiple plots side-by-side (or alt. row-by-row)
ggpubr::ggarrange(
  trust.likert.treat.az,
  trust.likert.treat.lcl,
  ncol = 2,
  common.legend = T,
  legend = "bottom", 
  font.label = list(family = "Palatino Linotype")
) +
  # add figure footnote annotation
  patchwork::plot_annotation(
    caption = str_wrap("Missing values removed. Displays percent proportions of responses by experiment condition and combined percentage proportions for two categories.",
               width = 85), theme = theme(text = element_text(family = "Palatino Linotype")))

```{r}
#| label: fig-distrustlikert
#| fig-cap: "Disrust in Election Administration by Experiment Condition"
#| fig-height: 4.5
#| fig-pos: "h"

# distrust in elections in AZ by experimental condition
distrust.likert.az <- df |>
  # reverse levels of factor vars to show positive difference in "not likely"
  mutate(across(c(q28_1:q28_5), ~forcats::fct_rev(.))) |> 
  ggstats::gglikert(
    include = c(q28_1:q28_5),
    y = 'group',
    facet_rows = vars(.question),
    variable_labels = c(
      q28_1 = "Voter fraud will occur",
      q28_2 = "Many votes won't be counted",
      q28_3 = "Many voters will be turned away",
      q28_4 = "Foreign Interference",
      q28_5 = "Election officials will discourage voting"),
    facet_label_wrap = 10,
    labels_size = 2.5,
    symmetric = F,
    add_totals = T
  ) +
  ggplot2::scale_fill_grey(start = 0.5, end = 0.1)+
  # ggplot2::scale_fill_brewer(palette = "Greys")+
  labs(subtitle = str_wrap("Maricopa County, Arizona", width = 95)) +
  theme_bw() +
  theme(
    text = element_text(family = "Palatino Linotype"),
    legend.position = 'bottom',
    axis.text.x = element_blank(), # remove percentage text along x-axis
    strip.text.y.right = element_blank()
  )

# distrust in elections in Local area by experimental condition
distrust.likert.lcl <- df |>
  # reverse levels of factor vars to show positive difference in "not likely"
  mutate(across(c(q40_1:q40_5), ~forcats::fct_rev(.))) |> 
  ggstats::gglikert(
    include = c(q40_1:q40_5),
    y = 'group',
    facet_rows = vars(.question),
    variable_labels = c(
      q40_1 = "Voter Fraud will occur",
      q40_2 = "Many Votes won't be counted",
      q40_3 = "Many voters will be turned away",
      q40_4 = "Foreign Interference",
      q40_5 = "Election officials will discourage voting"),
    facet_label_wrap = 10,
    labels_size = 2.5,
    symmetric = F,
    add_totals = T
  )+
  ggplot2::scale_fill_grey(start = 0.5, end = 0.1)+
  # ggplot2::scale_fill_brewer(palette = "Greys")+
  labs(subtitle = str_wrap("Local Area", width = 95)) +
  theme_bw() +
  theme(
    text = element_text(family = "Palatino Linotype"),
    legend.position = 'bottom',
    axis.text.y.left = element_blank(), # remove y-axis text on 2nd plot
    axis.text.x = element_blank(), # remove percentage text along x-axis
    strip.text.y.right = element_text(angle = 0, size = 8)
  )

# Instead of two separate plots, arrange into one plot while remaining separated
# use `ggpubr` package to print multiple plots side-by-side (or alt. row-by-row)
ggpubr::ggarrange(
  distrust.likert.az,
  distrust.likert.lcl,
  ncol = 2,
  common.legend = T,
  legend = "bottom", 
  font.label = list(family = "Palatino Linotype")
) +
  # add figure footnote annotation
  patchwork::plot_annotation(
    caption = str_wrap("Missing values removed. Displays percent proportions of responses by experiment condition and combined percentage proportions for two categories.",
               width = 85), theme = theme(text = element_text(family = "Palatino Linotype")))
